Extracting structured knowledge from document corpora --- turning unstructured text into typed, evidence-backed knowledge graphs --- remains a fundamental challenge across domains. Traditional approaches require domain-specific models retrained for each application, limiting generalizability. I present Epistract, an open-source framework that has evolved from a biomedical literature analysis tool into a domain-agnostic platform for agentic knowledge graph construction. The framework employs a three-layer architecture: a domain-agnostic core pipeline (document ingestion, parallel LLM extraction agents, graph construction via sift-kg, and epistemic analysis dispatch), pluggable domain configurations (entity/relation schemas, extraction prompts, and domain-specific epistemic rules defined in YAML and Python), and example consumers (web dashboard, Telegram bot) that work with any domain's graph output. Atop the domain-specific brute facts layer, a two-layer knowledge graph architecture adds an epistemic super-domain that surfaces hypotheses, contradictions, gaps, and risks --- analysis that transfers unchanged across domains. Originally developed for biomedical literature, the framework was validated across six drug discovery scenarios --- neurogenetics, oncology, rare disease, immuno-oncology, cardiovascular inflammation, and GLP-1 competitive intelligence --- extracting 783 nodes and over 2,200 links from 111 documents with 100\% entity type coverage and 93\% relation type coverage. To validate cross-domain generalization, I applied the same pipeline infrastructure to a corpus of 62 event planning contracts (PDFs, spreadsheets, and email correspondence), extracting 341 nodes and 663 edges with 53 cross-contract conflicts detected --- requiring only a new domain configuration package, with no modifications to the core pipeline. Adding a new domain requires defining a schema and extraction prompt; a domain creation wizard automates this process. The system is available as an MIT-licensed plugin installable in a single command.\footnote{\url{https://github.com/usathyan/epistract}}
