\section{Motivation: the missing dimension}
\label{sec:motivation}

Modern knowledge graph (KG) construction has converged on a recognizable pattern. Documents enter; an extractor (whether rule-based, fine-tuned NER, or LLM-driven) emits entities and typed relations; a graph builder deduplicates and resolves; a visualizer renders. The output is a labeled property graph that answers structural questions: which entities are most central, which clusters form, which relations are most prevalent. The flagship knowledge-graph platforms -- Neo4j Bloom, Stardog, GraphDB, Microsoft's GraphRAG~\cite{edge2024graphrag}, and a growing ecosystem of open-source LLM-based extractors -- all deliver this primary product.

A flat knowledge graph, however, is fundamentally lossy with respect to the epistemic content of its source corpus. Consider three sentences that might appear in a drug-discovery corpus:

\begin{enumerate}
\item ``In SURPASS-2, tirzepatide 15 mg reduced HbA1c by 2.30\% versus 1.86\% for semaglutide 1 mg (p<0.001).'' \cite{frias2021surpass2}
\item ``In one or more embodiments, the compositions of the present invention may be useful for the treatment of cardiovascular disease, neurodegenerative disorders, and metabolic dysfunction.''
\item ``GLP-1 receptor agonism has been hypothesized to confer neuroprotective effects through mechanisms that are not yet fully characterized.''
\end{enumerate}

All three sentences are about drug-disease relationships, and a typed-relation extractor (the dominant pattern in modern KG construction) will produce structurally similar edges from each. To make the comparison concrete, take sentence~(1). A traditional flat KG -- whether built by Neo4j-with-spaCy, GraphRAG, or any of the open-source LLM extractors -- would yield approximately the following from this single sentence:

\begin{quote}\small
\textbf{Entities.} \texttt{tirzepatide} (Compound), \texttt{semaglutide} (Compound), \texttt{HbA1c} (Biomarker), \texttt{SURPASS-2} (ClinicalTrial), \texttt{type 2 diabetes} (Disease, often inferred from HbA1c context).\\[2pt]
\textbf{Typed relations.}\\
\hspace*{1em}\texttt{tirzepatide REDUCES HbA1c} \quad (conf 0.95)\\
\hspace*{1em}\texttt{semaglutide REDUCES HbA1c} \quad (conf 0.95)\\
\hspace*{1em}\texttt{tirzepatide EVALUATED\_IN SURPASS-2} \quad (conf 0.99)\\
\hspace*{1em}\texttt{semaglutide EVALUATED\_IN SURPASS-2} \quad (conf 0.99)\\
\hspace*{1em}\texttt{tirzepatide INDICATED\_FOR type\_2\_diabetes} \quad (conf 0.85)\\
\hspace*{1em}\texttt{semaglutide INDICATED\_FOR type\_2\_diabetes} \quad (conf 0.85)\\[2pt]
\textbf{Edge metadata available.} Confidence score; verbatim evidence quote; source document identifier.
\end{quote}

This is a perfectly serviceable labeled property graph. It is typed and directional, it carries source attribution, and it is queryable: ``which compounds are indicated for type-2 diabetes?'' returns both. What it cannot tell the analyst is that this answer comes from a Phase 3 head-to-head with $p<0.001$. Querying the same flat graph after sentence~(2) is also ingested would return identical-shape edges asserting tirzepatide is indicated for cardiovascular disease and neurodegenerative disorders -- claims that are patent-prophecy, not empirically established. Confidence is a poor discriminator: a thorough patent extractor will produce \texttt{prophetic} edges with high confidence (the patent really does claim them), and a hedged review extractor will produce \texttt{hypothesized} edges with lower confidence -- but the confidence axis collapses the categorical distinction. Sentence~(2) yields edges with status \emph{prophetic} (rule: ``in one or more embodiments\ldots'' phrase + document-type-is-patent inference); sentence~(3) yields \emph{hypothesized} (rule: ``has been hypothesized'' regex). Two graphs containing nominally identical edges produce different decisions for an analyst reading them.

This is not a problem that domain-specific extraction prompting solves. The information is present in the source text -- the syntax of patents, hedging vocabulary in research papers, the structural conventions of trial reports -- but a graph that records only $\langle$ subject, predicate, object, confidence, source $\rangle$ has no field in which to store it. The missing dimension is \emph{epistemic typing}: a categorical label, attached to each relation, that records what kind of claim it is.

\paragraph{Who the framework is for, and what it is for.} Claude Code already provides rich \emph{deep-research} capabilities: built-in web search and document fetch, plus a growing ecosystem of plugins and MCP servers for specialized research workflows -- academic literature search, patent and regulatory database access, trial-registry retrieval, competitor and market intelligence, browser automation, third-party data stores. A scientist or computational analyst using Claude Code can already conduct deep research on a topic: collect papers, fetch patents, pull trial protocols, scrape regulatory filings. Epistract is positioned \emph{downstream} of that deep-research stage. Once a body of documents has been assembled -- through Claude Code's deep-research tools, manual curation, an institutional document store, or any combination -- Epistract is the tool that consolidates that body into a structured knowledge graph project: typed entities and relations grounded in the domain's ontology, categorical epistemic statuses on every relation, an analyst briefing, and a reactive chat surface. The graph project then becomes the substrate for answering many specific questions about the consolidated body. The framework is built for biomedical researchers, clinical-trial analysts, regulatory affairs teams, competitive-intelligence professionals, contract analysts, and others whose work involves rigorous synthesis of literature sets they have already assembled or are continuing to assemble through deep research.

It is primarily an \emph{analytical} tool: it does not transact, does not write back to the source corpus, and does not generate new documents. It produces a knowledge graph, an epistemic classification of every relation in that graph, and an analyst briefing. The scientist consumes those artifacts and asks follow-up questions in the workbench chat. The framework's job is to accelerate the synthesis-and-analysis phases of scientific work -- consolidating evidence, cataloging claims by epistemic status, surfacing contradictions and gaps, supporting follow-up question-answering -- while keeping deep research and analytical decision-making in the scientist's hands.

\paragraph{Bespoke consolidation, per question-bundle.} The graph that Epistract produces is \emph{bespoke}: it is shaped by the corpus a scientist assembled to answer a question or a bundle of related questions. ``How does the contemporary GLP-1 receptor agonist competitive landscape look across Phase 3 outcomes?'' is one such question-bundle -- and the deep research that answers it might span twenty-four PubMed papers, ten patent filings, and ten ClinicalTrials.gov protocols, assembled over hours or days. Epistract's role begins after that body is in hand. The framework consolidates the body into a graph whose entity types come from the chosen domain's schema (drug-discovery, clinicaltrials, or a domain the scientist extends via \texttt{/epistract:domain}), whose relations carry both confidence and categorical epistemic status, and whose persona-driven narrator surfaces the patterns -- prophetic concentrations, contested edges, gap signals -- that an analyst would otherwise have to find by hand. The same body, a different question-bundle, would call for a different consolidation pass and possibly a different domain.

\paragraph{Iterative refinement.} A scientist's understanding of a consolidated body is not a one-shot phenomenon. As they read the briefing and explore the graph, they discover gaps (a missing study, an unindexed patent), refinements to the analytical frame (a new entity type to track, a hedging pattern the rules missed, a contradiction worth promoting to a custom rule), and follow-up questions worth chasing. Some of those follow-ups send the scientist back to deep research -- using Claude Code's web tools or MCP-mediated source connectors to acquire missing documents -- and back to Epistract for reconsolidation. The framework supports this loop as discrete commands: \texttt{/epistract:ingest <new-folder>} appends documents to the corpus and re-extracts; \texttt{/epistract:epistemic} re-runs Layer 2 classification and the narrator briefing against the updated graph; the workbench chat reactively answers follow-ups against the current graph state. The framework itself does not auto-learn across runs -- refinements are human-mediated, and GitHub Issue~\#15 tracks the aspirational compounding mechanism. What does persist across runs is the \emph{scientist's} broader context: Claude Code's built-in memory system and its plugin / MCP ecosystem keep the accumulated knowledge -- what they have tried, what they have found, what they want to try next -- available across sessions. Deep research lives upstream of the consolidation; the consolidation feeds new questions; new questions drive more deep research. Epistract is the consolidation layer in that loop.

\paragraph{Returning to the architecture.} This paper introduces Epistract, a domain-agnostic knowledge graph framework whose primary contribution is to treat the epistemic dimension as a first-class architectural layer rather than as ad-hoc per-relation metadata. Epistract produces \emph{two-layer} graphs: Layer 1 is the conventional brute-facts graph (entities, typed relations, evidence quotes, confidence scores -- exactly what the worked example above shows the flat KG producing), and Layer 2 is an \emph{epistemic super-domain layer} that annotates every relation with a categorical status using a rule engine combining regex patterns over the verbatim evidence text with document-type inference. The two layers are produced in the same pipeline, persisted in the same artifact (\texttt{claims\_layer.json}), and rendered together in the visualization. A separate analyst-persona stage consumes the classified graph (not the source text) to produce a proactive briefing -- contradictions surfaced, prophetic claims separated from asserted ones, follow-up actions recommended.

Two design choices distinguish the framework from prior work. The first is \emph{domain pluggability}: the pipeline is domain-agnostic, and a domain is defined by four small configuration files (schema in YAML, extraction prompt in Markdown, epistemic rules in Python, analyst persona in YAML). The framework ships with four pre-built domains -- drug-discovery, contracts, clinicaltrials, and fda-product-labels -- and adding a new one requires no Python changes to the core pipeline. The second is the \emph{persona dual-use pattern}: a single YAML string per domain (the analyst persona) is the system prompt for both a reactive chat surface (workbench, where users ask ad-hoc questions) and a proactive briefing surface (the auto-generated narrator output). The two surfaces stay coherent because they derive from the same persona.

The remainder of this paper is organized as follows. Section~\ref{sec:architecture} describes the two-layer architecture and the pipeline. Section~\ref{sec:epistemic} details the epistemic super-domain layer -- the rule engine, the status taxonomy, and the persona-driven briefing. Section~\ref{sec:domains} describes the four pre-built domains and the \texttt{/epistract:domain} wizard for adding new ones. Section~\ref{sec:evaluation} reports validation across the eight scenarios spanning three active domains. Section~\ref{sec:evolution} situates the framework relative to similarity-based KG construction. Sections~\ref{sec:collaboration}--\ref{sec:conclusion} discuss collaboration, availability, and limitations.
